---
title: "milestone2"
author: "CHIA-YUN KUO"
date: "11/22/2021"
output:
  html_document:
    toc: true
    toc_depth: 2
---

## Install Required packages


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data

We import `infiltrating` ductular carcinoma and `lobular` Carcinoma htseq counts files.

```{r}
directory='./data/'

sampleFiles = grep("htseq.counts$", list.files('./data/', recursive = TRUE),value=TRUE)
length(sampleFiles)
```

`806` htseq counts file been loaded
but the number is incorrect, with further inspection, we found that the matching name pattern include 
1. *htseq.counts
2. *htseq_counts.txt
By modification `grep` pattern to `"htseq.counts$|htseq_counts"`, all the htseq counts files, `740 infiltrating` and `135 lobular` been loaded.

```{r}
sampleFiles = grep("htseq.counts$|htseq_counts", list.files('./data/', recursive = TRUE),value=TRUE)
length(sampleFiles)

sampleNames <-sub(".*./(.*).htseq.counts.*","\\1",sampleFiles)
sampleCondition <- sub("(.*.)/.*./.*.htseq.counts.*","\\1",sampleFiles)
barplot(table(sampleCondition))

sampleTable <- data.frame(sampleName = sampleFiles,
                          fileName = sampleFiles,
                          condition = sampleCondition)
sampleTable$condition <- factor(sampleTable$condition)

```

``` {r}
library("DESeq2")
ddsHTSeq <- DESeqDataSetFromHTSeqCount(sampleTable = sampleTable,
                                       directory = directory,
                                       design= ~ condition)
ddsHTSeq
```

pre-filtering reads count less than 9
``` {r}
dds <- ddsHTSeq[ rowSums(counts(ddsHTSeq)) > 9, ]
dds$condition <- factor(dds$condition, levels = c("infiltrating","lobular"))
dds$condition <- droplevels(dds$condition)
```

run Deseq2 and preview the result

``` {r}
dds <- DESeq(dds)
res <- results(dds)
res <- results(dds, contrast=c("condition","infiltrating","lobular"))
res
```

## Log fold change shrinkage for visualization and ranking

first preview the result name
``` {r}
resultsNames(dds)
```

than do the `LFC estimates`
``` {r}
resLFC <- lfcShrink(dds, coef="condition_lobular_vs_infiltrating", type="apeglm")
resLFC
```

## p-values and adjusted p-values

Order result table on *P value* in ascending order

``` {r}
resOrdered <- res[order(res$pvalue),]
summary(res)
```

# Exploring and exporting results

## MA-plot

``` {r}
plotMA(res, ylim=c(-2,2))
```


It is more useful visualize the MA-plot for the shrunken log2 fold changes, which remove the noise associated with log2 fold changes from low count genes without requiring arbitrary filtering thresholds.

``` {r}
plotMA(resLFC, ylim=c(-2,2))
```

## Alternative shrinkage estimators


``` {r}
# because we are interested in treated vs untreated, we set 'coef=2'
resNorm <- lfcShrink(dds, coef=2, type="normal")
resAsh <- lfcShrink(dds, coef=2, type="ashr")

```
